---
editor: visual
---

# Data, variables and methods

## Data and sample

Database used is the International Computer and Information Literacy (ICILS from now) study made by the International Association for the Evaluation of Educational Achievement (IEA) in 2018. The participants are students of eighth grade which were sampled using a random multistage stratified cluster design to be representative at the school and student level in all the countries measured. Schools can be referred to as the primary sampling units, and and students as the secondary sampling units.

The analysis focuses exclusively on the Chilean subset of the ICILS 2018. The sample comprised 3,092 students (1519 boys and 1573 girls) distributed across 178 schools. Both levels are representative of the broader Chilean student and school population.

The students underwent a Computer Literacy test and completed a questionnaire about their home background, values, beliefs, attitudes, and behaviors related to Computer and Information technologies. Also the principals and ICT managers of the school were surveyed to obtain information on the school's social characteristics and ICT infrastructure.

## Variables

### Dependent variables

**General and Specialized ICT Self-efficacy**: ICILS student data encompasses two indices of digital self-efficacy: one for general or basic applications and other for advanced or specialized tasks. Both indices were constructed using a same battery with the question *How well can you do this tasks when using ICT?*, and the answers 'I know how to do this' (1), 'I have never done this but i could work out how to do this (2)', 'I do not think I could do this' (3). The items can be examined in @tbl-selfeff-items[^1].

[^1]: Due to the extended length of the labels for items related to ICT self-efficacy, we truncated them in plots presentation. @tbl-labels illustrates both the original and the summarized labels for the items comprising both indexes and includes missing values n and missing values proportion.

```{r, results='asis', echo=FALSE}
#| label: tbl-selfeff-items
#| tbl-cap: "ICT Self-efficacy items"
#| tbl-subcap: 
#|  - "General ICT Self efficacy"
#|  - "Specialized ICT Self efficacy"
#| layout-ncol: 1

general_items <- data |>
  select(is2g27a,is2g27c,is2g27d,is2g27i,
         is2g27j,is2g27k,is2g27l,is2g27m,)|>
  to_label()

specialized_items <- data|>
  select(is2g27b,is2g27e,is2g27g,is2g27h)|>
  to_label()

general_tbl <- dfSummary(general_items,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = T,    # missing
               graph.col = F, # plot
               valid.col = T # n valido
               )

specialized_tbl <- dfSummary(specialized_items,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = T,    # missing
               graph.col = F, # plot
               valid.col = T # n valido
               )

general_tbl$Variable <- NULL # delete variable column
general_tbl$`Stats / Values` <- NULL
general_tbl$`Freqs (% of Valid)` <- NULL

specialized_tbl$Variable <- NULL # delete variable column
specialized_tbl$`Stats / Values` <- NULL
specialized_tbl$`Freqs (% of Valid)` <- NULL

general_tbl|>
  kable()
specialized_tbl|>
  kable()
```

The construction of both self-efficacy indices suggested by IEA were validated at international level, with a comparative approach between countries. A confirmatory factor analysis between the items above was carried out to validate the index at chilean sample scale[^2]. Both indexes were constructed by the IEA using the Weighted Likelihood Estimation Method, where scores have a mean of 50 and a standard deviation of 10, based on equal weights for all countries. A higher index value indicates a higher level of self-efficacy.

[^2]: See @fa-analysis for more context

```{r}
#| label: tbl-descriptives-selfeff
#| tbl-cap: Descriptives dependent variables
  
efficacies <- data|>
  select(s_geneff,s_speceff)

efficacies_tbl <- dfSummary(efficacies,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T) # n valido

efficacies_tbl$Variable <- NULL
efficacies_tbl$`Freqs (% of Valid)` <- NULL


efficacies_tbl|>
  kable()
```

### Independent variables

***Student level***

**Student sex**: Provided in the ICILS database as dichotomic, where Girl is 1 and Boy 0 [^3].

[^3]: We speak of student sex rather than gender because the ICILS study presents it as such.

**Computer and Information Literacy**: CIL scale is based on a computer application test with multi-answer questions and tasks based on a realistic theme and following a linear narrative structure. The test measure Computer Abilities as website construction, digital files managing database building, mapping tools, social media posting and internet search of information. CIL reporting scale was established with a mean of 500 (the average CIL scale score across countries) and a standard deviation of 100 for the equally weighted national samples that met ICILS 2013 (The first cycle of the study). Due to the large magnitude of this variable, the scale had to be multiplied by 0.1 to make the results more intelligible. The higher the value of the scale, the higher the CIL of the student [^4].

[^4]: ICILS database offers five possible values of the CIL score that were generated with full conditioning to derive summary student achievement statistics. Conventionally, papers based on this study usually occupy the first plausible value, which is coded "pv1cil". This research is no exception. For more information, read the ICILS 2018 User Guide.

**Control variables**: The analysis includes control variables that tend to be influential in studies of educational and technological inequalities. These variables encompass family socioeconomic status - by the highest level of parental education (ISCED Scale) - and the number of books in the household.

@tbl-individual-descriptives displays key statistics of independent student-level vairables to explore.

```{r, results='asis', echo=FALSE}
#| label: tbl-individual-descriptives
#| tbl-cap: "Individual Variables Descriptives"

independent_student <- data|>
  select(
    s_sex,s_pv1cil,
    s_hisced,s_homlit
  )|>
  mutate(
    s_sex=to_label(s_sex),
    s_hisced=to_label(s_hisced),
    s_homlit=to_label(s_homlit)
  )

tbl_individual<-dfSummary(independent_student,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

tbl_individual$Variable <- NULL # delete variable column

print(tbl_individual)
```

***School level***

The study took into account two contextual variables to develop multilevel analyses. The first is the average score on the Computer and Information Literacy test at school, which was estimated by aggregating students' scores by mean. The second is the gender composition of the school, which is estimated by dividing the total number of female students by the total number of eighth grade students in the school. According to the resulting ratio of female students, schools were classified into three categories. @tbl-school-descriptives details frequencies and descriptive statistics of these variables.

```{r, results='asis', echo=FALSE}
#| label: tbl-school-descriptives
#| tbl-cap: "School Variables Descriptives"

independent_school <- data|>
  mutate(c_gender_type=to_label(c_gender_type))|>
  distinct(idschool, .keep_all = TRUE)|>
  select(c_pv1cil,c_gender_type)

tbl_school<-dfSummary(independent_school,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

tbl_school$Variable <- NULL # delete variable column

print(tbl_school)
```

## Method

The analysis start deploying bivariate and correlation exploration of the data, reaching gender differences in self-efficacy distribution and their determinants. Then, due to the hierarchical structure of the data, as well as its representativeness at the student and school level, linear regression models are estimated from a multilevel perspective. This means that the models are oriented towards estimating both individual and contextual effects in a single analysis; capturing variation between groups and within groups [@vijverMultilevelAnalysisIndividuals2008; @hoxMultilevelAnalysisTechniques2017].

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://osf.io/5qjsd). The statistical analysis of this research was performed using the free software R version 4.3.0.
