---
editor: visual
bibliography: "icils_nudos.bib"
---

```{r setup two}
options(scipen=999) #No scientific notation
options(digits = 2) #Round results two decimals

pacman::p_load(tidyverse, #Data manipulation
               gt, #Tables 
               gtsummary, #Descriptive table
               sjmisc, #Label manipulation
               labelled, #Label manipulation
               skimr, #Data manipulation
               sjPlot, #Correlation
               corrplot, #Correlation
               Hmisc, #Correlation
               lme4, #Multilevel modeling
               reghelper, #Get ICC
               texreg, #View models
               AICcmodavg, #Best fit model
               responsePatterns, #Check items
               ggrepel, #labels plots
               ggpubr, #arrange plots
               ggExtra, #marginal plots
               scales,
               summarytools, #descriptive tables
               knitr, #Markdown
               kableExtra, #Markdown
               psych, #FA Analysis
               car,
               haven, #Open data
               janitor, #Data process
               Hmisc, #Labels manipulation
               stringr #Label manipulation
               )

# Open student and school raw data
bsgchli2 <- read_sav("../../../input/raw_data/2018/BSGCHLI2.sav")
bcgchli2 <- read_sav("../../../input/raw_data/2018/BCGCHLI2.sav") 

source("00_proccess.R")
```

# Results

The analysys begins shwing descriptive results for the self-efficacy items, as displayed in @fig-barplot-general. Notably, across all tasks, a significant proportion of students ---no less than 65%--- expressed confidence in their ability to perform general tasks, while fewer than 10% reported an inability to do so. On the whole, participants demonstrated a high level of general self-efficacy. In contrast, when it comes to specialized self-efficacy, less than 30% of students confidently state their proficiency in tasks such as building a webpage or creating a computer program. This percentage increases a little when it comes to tasks like setting up a local area network (31%) or creating a database (35%). Therefore, and as it could be expected, students tend to perceive having less skills in advanced than in basic ICT tasks.

```{r}
#| label: fig-barplot-general
#| fig-cap: "General and specialized ICT self-efficacy. Data: ICILS Chile 2018"

selfeff_items <- data|>
  select(is2g27a,is2g27c,is2g27d,is2g27i,
         is2g27j,is2g27k,is2g27l,is2g27m,
         is2g27b,is2g27e,is2g27g,is2g27h)|>
  to_label()

selfeff_items <- to_label(selfeff_items)

relative_freq_selfeff <- rbind(
  selfeff_items |>
    drop_na(is2g27a) |>
    count(variable = is2g27a) |>
    mutate(prop = n / sum(n), name = "Edit graphic images",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27c) |>
    count(variable = is2g27c) |>
    mutate(prop = n / sum(n), name = "Write or edit text",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27d) |>
    count(variable = is2g27d) |>
    mutate(prop = n / sum(n), name = "Search and find information on internet",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27i) |>
    count(variable = is2g27i) |>
    mutate(prop = n / sum(n), name = "Create multi-media presentation",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27j) |>
    count(variable = is2g27j) |>
    mutate(prop = n / sum(n), name = "Upload multimedia to an online profile",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27k) |>
    count(variable = is2g27k) |>
    mutate(prop = n / sum(n), name = "Insert an image into a document/message",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27l) |>
    count(variable = is2g27l) |>
    mutate(prop = n / sum(n), name = "Install a program/app",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27m) |>
    count(variable = is2g27m) |>
    mutate(prop = n / sum(n), name = "Judge internet information veracity",
           type = "General Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27b) |>
    count(variable = is2g27b) |>
    mutate(prop = n / sum(n), name = "Create a database",
           type = "Specialized Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27e) |>
    count(variable = is2g27e) |>
    mutate(prop = n / sum(n), name = "Build a webpage",
           type = "Specialized Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27g) |>
    count(variable = is2g27g) |>
    mutate(prop = n / sum(n), name = "Create a computer program/app",
           type = "Specialized Self-efficacy"),
  
  selfeff_items |>
    drop_na(is2g27h) |>
    count(variable = is2g27h) |>
    mutate(prop = n / sum(n), name = "Set up a local area network",
           type = "Specialized Self-efficacy")
)

relative_freq_selfeff <- relative_freq_selfeff |>
  mutate(
    name = factor(name, levels = c(
      "Write or edit text",
      "Search and find information on internet",
      "Insert an image into a document/message",
      "Install a program/app",
      "Upload multimedia to an online profile",
      "Edit graphic images",
      "Judge internet information veracity",
      "Create multi-media presentation",
      "Create a database",
      "Set up a local area network",
      "Build a webpage",
      "Create a computer program/app")),
    variable = factor(variable, levels = c(
      "I know how to do this.",
      "I have never done this but I could work out how to do this.",
      "I do not think I could do this."
    )),
    prop = round(prop, 2)
  )

relative_freq_selfeff |>
  ggplot(aes(x = prop, y = name, fill = variable)) +
  geom_bar(position = "fill", stat = "identity", color = "black") +
  facet_wrap(~type,ncol=1,scales = "free_y") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 42)) +
  labs(
    title = "How well can you do each of these tasks when using ICT?",
    x = "", y = "", fill = "",
    caption = "Missing values where deleted to calculate proportions"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.spacing = unit(1, "lines")
  ) +
  scale_fill_grey(
    start = 0.5, end = 1,
    limits = c(
      "I do not think I could do this.",
      "I have never done this but I could work out how to do this.",
      "I know how to do this."
    ),
    labels = ~ stringr::str_wrap(.x, width = 30)
  ) +
  scale_x_continuous(labels = scales::percent) +
  geom_text(aes(label = paste0(prop * 100, "%")), position = position_fill(vjust = 0.5))
```

Turning now to geneder differences in self-efficacy, @fig-cleveland-gender presents a comparison of male and female students, displaying the proportion of those answering "I know how to do this" in each gender group. Whereas girls outperform boys in most of the general self-efficacy items, the opposite occurs for the specialized one. While most of the differences are statistically significant, they tend to be larger for the specialized self-efficacy in favor of boys.

```{r}
summary_stats <- data_model %>%
  group_by(s_sex) %>%
  summarise(
    mean_pv1cil = mean(s_pv1cil, na.rm = TRUE),
    se_pv1cil = sd(s_pv1cil, na.rm = TRUE) / sqrt(n())
  )

# Create a bar plot with error bars
ggplot(summary_stats, aes(x = s_sex, y = mean_pv1cil, fill = s_sex)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_pv1cil - se_pv1cil, ymax = mean_pv1cil + se_pv1cil), width = 0.2) +
  labs(x = "Sex", y = "Mean of s_pv1cil", title = "Comparison of Means with Error Bars") +
  theme_minimal()


ggplot(data_model, aes(x = s_sex, y = s_pv1cil, fill = s_sex)) +
  geom_boxplot() +
  labs(x = "Sex", y = "s_pv1cil", title = "Box Plot of s_pv1cil by Sex") +
  theme_minimal()

# Perform t-test
t_test_result <- t.test(s_pv1cil ~ s_sex, data = data_model)

# Extract p-value and test statistic
p_value <- t_test_result$p.value
t_statistic <- t_test_result$statistic

library(ggplot2)
library(dplyr)

# Create the box plot
ggplot(data_model, aes(x = s_sex, y = s_pv1cil, fill = s_sex)) +
  geom_boxplot() +
  labs(x = "Sex", y = "s_pv1cil", title = "Information and Computer Literacy Test by Sex") +
  theme_minimal() +
  
  # Add annotation for the t-test results
  annotate("text", x = 2, y = min(data_model$s_pv1cil, na.rm = TRUE), 
           label = paste("t =", round(t_statistic, 2), "\n", "p =", round(p_value, 4)),
           size = 5, hjust = -0.2, vjust = -1, color = "red")




```

```{r}
#| label: fig-cleveland-gender
#| fig-cap: "General and specialized ICT Self-efficacy by gender. Data: ICILS Chile 2018"

data |>
  mutate(s_sex=to_label(s_sex))|>
  mutate_at(vars(starts_with("is2g")), ~ 
            ifelse(.==1,1,0)) |> #Dummy
  rename(gender=s_sex)|>
  group_by(gender) |>
  summarise(across(starts_with("is2g"),
                 mean, na.rm = TRUE)) |>
  pivot_longer(cols = starts_with(c("is2g")),
               names_to = "items",
               values_to = "puntaje") |>
  mutate(type=ifelse(items%in%c("is2g27b","is2g27e",
                                "is2g27g","is2g27h"),
                                "Specialized ICT self-efficacy",
                                "General ICT self-efficacy")) |>

  mutate(items=
  factor(items,
  levels=c(
  "is2g27b",
  "is2g27h",
  "is2g27e",
  "is2g27g",
  "is2g27c",
  "is2g27d",
  "is2g27k",
  "is2g27l",
  "is2g27j",
  "is2g27a",
  "is2g27m",
  "is2g27i"),
  labels=c(
  "Create a database",
  "Set up a local area network***",
  "Build a webpage*",
  "Create a computer program/app***",
  "Write or edit text***",
  "Search and find information on internet*",
  "Insert an image into a document/message**",
  "Install a program/app",
  "Upload multimedia to an online profile**",
  "Edit graphic images***",
  "Judge internet information veracity***",
  "Create multi-media presentation**"
  )))|>

  ggplot(aes(x=to_label(puntaje),y=to_label(items)))+
    geom_line(aes(group=to_label(items)),color="#E7E7E7",linewidth=2.0)+
    geom_point(aes(shape=gender,color=gender),size=4)+
    theme_minimal()+
    labs(x="Proportion of 'I know how to do this'",y="",color="Gender",
         caption= "Signif. codes Chi-square test: <0.001 = ***, <0.01 = **, <0.05 = *")+
    scale_color_grey(start=.2,end = .6)+
    theme(legend.position = "top")+
    scale_y_discrete(labels=function(x) str_wrap(x,width = 50))+
    facet_wrap(~type,ncol=1,scales="free_y")+
    labs(title="How well can you do each of these tasks 
when using ICT?")+
    guides(shape=guide_legend(title="Gender"))
```

@fig-gender-school-mean shows differences in general and specialized self-efficacy according to the gender composition of the schools. Schools are classified here as masculinized when they have less than one third of girls, feminized when this proportion is more than two thirds, and mixed when the girls' proportion lies in between. Regarding general self-efficacy we observe that the outperformance of girls is larger in feminized schools, whereas the opposite occurs for specialized self-efficacy: girls underperform boys in feminized schools.

```{r}
#| label: fig-gender-school-mean
#| fig-cap: "ICT Self Efficacy and CIL by Student sex and School composition. Data: ICILS Chile 2018"

data|>
  mutate(s_sex = to_label(s_sex),
         c_gender_type = labelled(c_gender_type,
                             labels = c("Masculinized school" = 1, 
                                        "Mixed school" = 2, 
                                        "Feminized school" = 3)),
         c_gender_type = to_label(c_gender_type)
         )|>
  group_by(s_sex,c_gender_type)|>
  drop_na(c_gender_type)|>
  summarise(geneff = mean(s_geneff, na.rm = TRUE),
            speceff = mean(s_speceff, na.rm = TRUE))|>
  pivot_longer(cols = c(geneff, speceff), names_to = "efficacy", values_to = "mean")|>
  
  ggplot(aes(x = c_gender_type, y = mean)) +
  geom_point(aes(shape = s_sex, color = s_sex), size = 3.5)+
  facet_wrap(~efficacy,
             nrow = 2,
             labeller = labeller( efficacy = c(
               "geneff" = "General ICT Self efficacy",
               "speceff" = "Specialized ICT Self efficacy"))
             )+
  scale_color_grey(start=.2,end = .6)+
  theme_minimal()+
  labs(shape = "Gender", color = "Gender",
       y = "Mean score", x= "",
       title = "Gender CIL and ICT Self-efficacy by type of School")+
  theme(legend.position = "top")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

*Multilevel models*

@tbl-general-multilevel and @tbl-specialized-multilevel displays the results of the multilevel regression models. Beggining with the results for the general self-efficacy, we observe in Model 1 that girls obtain better scores, but this is no longer significant when entering the Computer and Information Literacy (CIL) score in Model 2, with a positive significant association throughout the models. Regarding level 2 predictors, it is relevant to mention that the variance of self-efficacy is related to the schools (intra-class correlation of the null model) is only 5%, therefore there is few variance left to find significant effects at this level. The only predictor that display a significant effect at this level is the average school CIL, which, contrary to level 1, has a negative association with generalized self efficacy. Interactions effects are not significant in this case.

```{r}
#| label: tbl-general-multilevel
#| tbl-cap: "General Self-efficacy Multilevel models"

pacman::p_load(sjlabelled)
get_label(data_model$s_geneff)

data_model$s_geneff <- set_label(data_model$s_geneff, label = "Basic self-efficacy")

data_model <- data|>
  mutate(s_sex = to_label(s_sex),
         c_gender_type = labelled(c_gender_type,
                             labels = c("Masculinized school" = 1, 
                                        "Mixed school" = 2, 
                                        "Feminized school" = 3)),
         c_gender_type = to_label(c_gender_type)



data_model <- data |> mutate(
  s_sex=to_label(s_sex),
  c_gender_type = labelled(c_gender_type,
                             labels = c("Masculinized school" = 1, 
                                        "Mixed school" = 2, 
                                        "Feminized school" = 3)),
  c_gender_type = to_label (c_gender_type),
  c_gender_type = relevel(c_gender_type, "Mixed school")
  ) 

#Null model
m0_geneff <- lmer(s_geneff ~ 1 + (1 | idschool), data=data_model)

#Level 1 fixed effects
m1_a_geneff  <- lmer(s_geneff ~ 1 +
                       s_sex +
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model) 

m1_b_geneff  <- lmer(s_geneff ~ 1 +
                       s_sex + s_pv1cil +
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model) 

m1_c_geneff <- lmer(s_geneff ~ 1 +
                        s_sex + s_pv1cil +
                        s_sex*s_pv1cil +
                        s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                      data=data_model) 


# Level 2 fixed effects
m2_a_geneff <- lmer(s_geneff ~ 1 +
                       s_sex + s_pv1cil +
                       c_gender_type +
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)


m2_b_geneff <- lmer(s_geneff ~ 1 +
                       s_sex + s_pv1cil +
                       c_pv1cil +
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)

m2_c_geneff <- lmer(s_geneff ~ 1 +
                       s_sex + s_pv1cil +
                       c_pv1cil+ c_gender_type +
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)

m2_d_geneff <- lmer(s_geneff ~ 1 +
                       s_sex + s_pv1cil +
                       c_pv1cil+ c_gender_type +
                       s_sex * c_gender_type + 
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)

m2_e_geneff <- lmer(s_geneff ~ 1 +
                       s_sex + s_pv1cil +
                       c_pv1cil+ c_gender_type +
                       s_sex * c_pv1cil + 
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)

tab_model(m1_a_geneff,m1_b_geneff,m1_c_geneff,
          m2_a_geneff,m2_b_geneff,m2_c_geneff,
          m2_d_geneff,m2_e_geneff,
          show.ci = FALSE, auto.label = FALSE,
          p.style = "stars",collapse.se = TRUE,
          show.re.var = TRUE,show.icc = FALSE,
          show.obs = TRUE,show.ngroups = TRUE,
          show.aic = TRUE,
          rm.terms = c("s_hisced","s_homlit"),
          pred.labels = c("Intercept",
                          "Gender (Girl = 1)",
                          "CIL score",
                          "Gender*CIL",
                          "Gender composition: Masculinized school",
                          "Gender composition: Feminized school",
                          "School CIL",
                          "Gender*Masculinized school",
                          "Gender*Feminized school",
                          "MGender*School CIL"),
          dv.labels = c("Model 1","Model 2","Model 3",
                        "Model 4", "Model 5", "Model 6", "Model 7","Model 8"))

```

Turning now to specialized self-efficacy, @tbl-specialized-multilevel follow the same structure than the previous table. In this case, and contrary to what was observed for general self-efficacy, girls consistently show a lower average score than boys, whereas the CIL score now is negative and its statistical significance dissappears when adding predictors. Despite having a low intra-class correlation (6%), in this case it is possible to detect some relevant effects in level 2 predictors. School CIL, as in general self-efficacy, displays a negative effect, and in Model 7 the gender composition appears to have a positive association with specialized self-efficacy for masculinized (with mixed schools as the reference category), but non-significant association for feminized school. Regarding interactions, effects are not significant.

```{r}
#| label: tbl-specialized-multilevel
#| tbl-cap: "Specialized Self-efficacy Multilevel models"


data_model$s_speceff <- set_label(data_model$s_speceff, label = "Advanced self-efficacy")

# Null model
m0_speceff <- lmer(s_speceff ~ 1 + (1 | idschool), data=data_model)

# Level 1 fixed effects
m1_a_speceff <- lmer(s_speceff ~ 1 +
                     s_sex +
                     s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                   data=data_model) 

m1_b_speceff <- lmer(s_speceff ~ 1 +
                     s_sex + s_pv1cil +
                     s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                   data=data_model) 

m1_c_speceff <- lmer(s_speceff ~ 1 +
                     s_sex + s_pv1cil +
                     s_sex*s_pv1cil +
                     s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                   data=data_model) 

# Level 2 fixed effects
m2_a_speceff <- lmer(s_speceff ~ 1 +
                     s_sex + s_pv1cil +
                     c_gender_type +
                     s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                   data=data_model)


m2_b_speceff <- lmer(s_speceff ~ 1 +
                     s_sex + s_pv1cil +
                     c_pv1cil +
                     s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                   data=data_model)

m2_c_speceff <- lmer(s_speceff ~ 1 +
                     s_sex + s_pv1cil +
                     c_pv1cil+ c_gender_type +
                     s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                   data=data_model)

m2_d_speceff <- lmer(s_speceff ~ 1 +
                       s_sex + s_pv1cil +
                       c_pv1cil+ c_gender_type +
                       s_sex * c_gender_type + 
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)

m2_e_speceff <- lmer(s_speceff ~ 1 +
                       s_sex + s_pv1cil +
                       c_pv1cil+ c_gender_type +
                       s_sex * c_pv1cil + 
                       s_hisced + s_homlit +
                       (1 + s_sex| idschool),  
                     data=data_model)

#Final table
tab_model(m1_a_speceff,m1_b_speceff,m1_c_speceff,
          m2_a_speceff,m2_b_speceff,m2_c_speceff,
          m2_d_speceff,m2_e_speceff,
          show.ci = FALSE, auto.label = FALSE,
          p.style = "stars",collapse.se = TRUE,
          show.re.var = TRUE,show.icc = FALSE,
          show.obs = TRUE,show.ngroups = TRUE,
          show.aic = TRUE,
          rm.terms = c("s_hisced","s_homlit"),
          pred.labels = c("Intercept",
                          "Gender (Girl = 1)",
                          "CIL score",
                          "Gender*CIL",
                          "Gender composition: Masculinized school",
                          "Gender composition: Feminized school",
                          "School CIL",
                          "Gender*Masculinized school",
                          "Gender*Feminized school",
                          "Gender*School CIL"),
          dv.labels = c("Model 1","Model 2","Model 3",
                        "Model 4", "Model 5", "Model 6",
                        "Model 7", "Model 8"))



sjPlot::plot_models(m2_d_geneff,m2_d_speceff,
                    wrap.labels = 20,
                    show.p = T,
                    show.values = T,
                    show.intercept = F,
                    colors = c("#5f5758","#ff3057")) +
  theme_minimal()

```
