---
title: "Factores explicativos de las brechas de autoeficacia tecnológica por género en el sistema educativo chileno"
author: NUDOS
date: 26-10-2023
bibliography: ../../input/bib/informe1.bib 
csl: ../../input/bib/apa.csl

format:
  html:
    embed-resources: true
    toc: true
    number-sections: false
    highlight-style: github
    code-fold: true
    comments: 
      hypothesis: true
            
execute: 
  error: false
  warning: false
  message: false
---

# Introduction

```{r}
#0. Preparation ----

options(digits = 2) #round results to 2 digits

nudos_color_binary <- c("#5f5758","#ff3057") #Nudos theme

pacman::p_load(tidyverse, #Data manipulation
               stargazer, #Descriptive tables
               sjmisc, #Descriptive tables
               knitr, #Markdown
               corrr, #Correlation matrix
               sjlabelled, #Labels manipulation
               sjPlot, #View_df table
               Hmisc, #Manipulate labels
               ggplot2, #Plots
               hrbrthemes, #Plot themes
               labelled #Manipulate labels
               )

data <- readRDS("../../input/proc_data/03_student_proc_2018.rds")

#Student level
student <- data %>% select(starts_with("s_"))

#School level
school <- data %>% select(starts_with("c_"),idschool) %>% group_by(idschool) %>% 
                  summarise(across(everything(),~ mean(.x, na.rm = TRUE))) %>% ungroup() %>%
                  select(-idschool)

```

## Index presentations

La medición del año 2018 tiene 4 tipos de indicadores que pueden ser útiles como predictores de autoeficacia:
  
-   Actitudes hacia las tecnologías.
-   Actividades digitales fuera de la escuela.    
-   Actividades digitales dentro de la escuela.
-   Puntaje de la prueba.

:::{.panel-tabset}

#### Student Level
```{r,results='asis'}
student%>% mutate(s_sex=as_factor(s_sex))%>%
  view_df(show.id = F, show.values = F, show.type = T,show.labels = F)
```


#### School Level
```{r,results='asis'}
label(school[["c_t_eff"]]) <- "School mean teacher ICT self-efficacy"
label(school[["c_t_vwpos"]]) <- "School mean teacher positive ICT views"
label(school[["c_t_vwneg"]]) <- "School mean teacher negative ICT views"
label(school[["c_speceff"]]) <- "School mean specific self-efficacy"
label(school[["c_geneff"]]) <- "School mean general self-efficacy"
label(school[["c_ictneg"]]) <- "School mean negative ICT percepctions"
label(school[["c_ictpos"]]) <- "School mean positive ICT perceptions"
label(school[["c_ictlrn"]]) <- "School mean learning tsks ICT at school"
label(school[["c_usestd"]]) <- "School mean use ICT study purposes"
label(school[["c_ictfut"]]) <- "School mean expectations ICT work-study"
label(school[["c_speclass"]]) <- "School mean specialist application ICT in class"
label(school[["c_genclass"]]) <- "School mean general applications ICT in class"
label(school[["c_codlrn"]]) <- "School mean Learning coding in class"
label(school[["c_pv1cil"]]) <- "School mean score CIL Test"
label(school[["c_f_ratio_grade"]]) <- "School proportion of females students in eigth grade"
label(school[["c_p_sex"]]) <- "Principal School sex"
label(school[["c_usecom"]]) <- "School mean use ICT social communication"
label(school[["c_useinf"]]) <- "School mean use ICT for information"
label(school[["c_accont"]]) <- "School mean use ICT accesing content"


school%>%
  view_df(show.id = F, show.values = F, show.type = T,show.labels = F)
```
:::

# Data distribution

## Descriptive summary
:::{.panel-tabset}

#### Student level
```{r, results='asis'}
student %>% 
  sjmisc::descr(show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
  kable()
```

#### School level

```{r, results='asis'}
sjmisc::descr(school%>%select(-c_accont,-c_p_sex),show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
kable("markdown")
```
:::

## Gap of second order
```{r}

student %>% 
  group_by(to_label(s_sex)) %>%
  summarise_at(c("s_geneff", "s_speceff",
                 "s_ictpos","s_ictneg","s_ictfut",
                 "s_usecom","s_useinf","s_accont"), 
                 mean, na.rm = TRUE) %>%

  rename(Gender="to_label(s_sex)") %>%
  pivot_longer(cols = starts_with("s_"),
               names_to="variables") %>%
  
  ggplot(aes(x=value,y=variables,color=Gender))+
  geom_point(size=6)+
  theme_ipsum()+
  labs(x="Value",y="Indicator",color="Gender")+
  scale_color_manual(values=nudos_color_binary)
  
```


## Self-efficacy and attitudes to ICT variance relations
```{r,results='asis'}
student %>% select(s_geneff, s_speceff,
                 s_ictpos,s_ictneg,s_ictfut,
                 s_usecom,s_useinf,s_accont)%>%
  tab_corr()
```



