---
title: "Factor Analyses: ICILS 2018 General and Specialized Self Efficacy"
author: "Daniel Miranda"

date: "`r Sys.Date()`"

output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    use_bookdown: true
pkgdown:
  as_is: true
  

---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
	error = TRUE,
	message = FALSE,
	warning = FALSE,
	echo= FALSE
)
```


```{r , echo=FALSE}
pacman::p_load(tidyverse,
               gtsummary,
               sjmisc, #Label manipulation
               labelled, #Label manipulation
               summarytools,
               sjPlot,
               effects,
               #responsePatterms
               stargazer, # Reporte a latex
                sjPlot, sjmisc, # reporte y gráficos
                corrplot, # grafico correlaciones
                xtable, # Reporte a latex
                Hmisc, # varias funciones
                psych, # fa y principal factors
                psy, # scree plot function
                nFactors, # parallel
                GPArotation, devtools, # rotación
               table1,
               xtable,
               apaTables, kableExtra
               )


student_proc_2018 <- readRDS("/Users/daniel/Dropbox (Personal)/github/multilevel_icils_gender_chile/input/proc_data/03_student_proc_2018.rds")
```

## Variables

How well can you do each of these tasks when using ICT? (Please mark one choice in each row)

Specialized ICT Self-efficacy

- Create a database (e.g., using [Microsoft Access®])	IS2G27B
 
- Build or edit a webpage	IS2G27E

- Create a computer program, macro, or [app] (e.g., in [Basic, Visual Basic])	IS2G27G

- Set up a local area network of computers or other ICT	IS2G27H


```{r}
#Plot items Specialized Self-efficacy
speceff_plot <- student_proc_2018 |>
  mutate(s_sex=to_label(s_sex),
         is2g27a=to_label(is2g27b),
         is2g27c=to_label(is2g27e),
         is2g27d=to_label(is2g27g),
         is2g27i=to_label(is2g27h))

m1=rbind(
  speceff_plot|>count(variable=is2g27b)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="b.Create a database"),
  speceff_plot|>count(variable=is2g27e)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="e.Build or edit a webpage"),
  speceff_plot|>count(variable=is2g27g)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="g.Create a computer program, macro, or [app]"),
  speceff_plot|>count(variable=is2g27h)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="h.Set up a local area network")
) |>
  mutate(prop=round(prop,2)) 
#View(m1)
#str(m1)
m1$variable=as.factor(m1$variable)

  ggplot(m1, aes(x=name,y=prop,fill=variable))+
 geom_bar(position = "fill",stat="identity",color="black")+
  scale_fill_manual(values = c("black","grey50","grey95"),
                    labels = c("I know how", "I never done but I could", "I think I couldn't" ))+
  coord_flip()+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25))+
  labs(title="Specialized Self-efficacy",
       subtitle="How well can you do each of these tasks when using ICT?",
       x="",y="",fill="")+
  theme(legend.position = "top")+
  guides(fill = guide_legend(reverse=TRUE))+
  scale_y_continuous(labels = scales::percent)
```


General ICT Self-efficacy

- Edit digital photographs or other graphic images	IS2G27A

- Write or edit text for a school assignment	IS2G27C

- Search for and find relevant information for a school project on the Internet	IS2G27D

- Create a multi-media presentation (with sound, pictures, or video)	IS2G27I

- Upload text, images, or video to an online profile	IS2G27J

- Insert an image into a document or message	IS2G27K

- Install a program or [app]	IS2G27L

- Judge whether you can trust information you find on the Internet	IS2G27M

```{r}
#Plot items General self-efficacy
geneff_plot <- student_proc_2018 |>
  mutate(s_sex=to_label(s_sex),
         is2g27a=to_label(is2g27a),
         is2g27c=to_label(is2g27c),
         is2g27d=to_label(is2g27d),
         is2g27i=to_label(is2g27i),
         is2g27j=to_label(is2g27j),
         is2g27k=to_label(is2g27k),
         is2g27l=to_label(is2g27l),
         is2g27m=to_label(is2g27m))


m=rbind(
  geneff_plot|>count(variable=is2g27a)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="a.Edit digital photographs"),
  geneff_plot|>count(variable=is2g27c)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="c.Write or edit text"),
  geneff_plot|>count(variable=is2g27d)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="d.Search for and find relevant information"),
  geneff_plot|>count(variable=is2g27i)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="i.Create a multi-media presentation"),
  geneff_plot|>count(variable=is2g27j)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="j.Upload text, images, or video"),
  geneff_plot|>count(variable=is2g27k)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="k.Insert an image"),
  geneff_plot|>count(variable=is2g27l)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="l.Install a program or [app]"),
  geneff_plot|>count(variable=is2g27m)|>drop_na()|>
    mutate(prop=n/sum(n),
           name="m.Judge whether you can trust information")
  ) |>
  mutate(prop=round(prop,2)) 
#View(m)

  ggplot(m, aes(x=name,y=prop,fill=variable))+
  geom_bar(position = "fill",stat="identity",color="black")+
  scale_fill_manual(values = c("black","grey50","grey95"),
                    labels = c("I know how", "I never done but I could", "I think I couldn't" ))+
  coord_flip()+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25))+
  labs(title="General Self-efficacy",
       subtitle="How well can you do each of these tasks when using ICT?",
       x="",y="",fill="")+
  theme(legend.position = "top")+
  guides(fill = guide_legend(reverse=TRUE))+
  scale_y_continuous(labels = scales::percent)
```



```{r , echo=FALSE}
base <- student_proc_2018 |> 
  select(#s_sex,
         #s_speceff,s_geneff,
         #Items specific self-efficacy
         is2g27b,is2g27e,is2g27g,is2g27h,
         #Items general self-efficacy
         is2g27a,is2g27c,is2g27d,is2g27i,is2g27j,is2g27k,is2g27l,is2g27m)

baserev <-base |> 
       mutate_at(vars(starts_with("is2g")),
                 ~4-.)

#variables
baserev=baserev |>
  mutate(database=is2g27b,
         webpage=is2g27e,
         app=is2g27g,
         network=is2g27h,
         editphoto=is2g27a,
         edittext=is2g27c,
         findinfo=is2g27d,
         multimedia=is2g27i,
         uploadfile=is2g27j,
         insertimage=is2g27k,
         install=is2g27l,
         judgeinfo=is2g27m
         ) |>
  select(database, webpage, app, network, editphoto, edittext, findinfo, multimedia, uploadfile, insertimage, install, judgeinfo)

```



## Correlation matrix: assuming continuous 

```{r}
#Test correlation (Now is positive between items and index, and negativa between items and cil)
mat = round(cor(baserev, use="complete.obs"),3)
#mat

#Hide lower triangle
lower<-mat
lower[lower.tri(mat, diag=TRUE)]<-""
lower<-as.data.frame(lower)

```

```{r, results='asis'}
#Table using kableExtra
lower |>
  kable() |>
#kbl(caption = "Recreating booktabs style table") |>
  kable_classic()

```

## Correlation matrix: assuming categorical

```{r}
#Polychoric correlations
m=psych::polychoric(baserev)
polmat=m$rho

#Hide lower triangle
lowerpol<-round(polmat, 3)
lowerpol[lower.tri(polmat, diag=TRUE)]<-""
lowerpol<-as.data.frame(lowerpol)


```

```{r}
#Table using kableExtra
lowerpol |>
kable() |>
  kable_classic_2()
```

```{r}
#Correlations graph
corrplot(polmat, type="lower") #lower x bajo diagonal

cor.plot(m$rho, numbers=T, upper=FALSE, main = "Polychoric Correlation", show.legend = FALSE)
```

## Paralel Analyses
```{r}
save(polmat, file = "polychoric")

# Factor analyses: 3 factors
load("polychoric")
fa.parallel(polmat, fm="pa", fa="fa", main = "Scree Plot", n.obs=3092)
```

## EFA considering four factors
```{r}
poly_model4 = fa(baserev, nfactor=4, cor="poly", fm="pa", rotate = "oblimin")

save(poly_model4, file = "poly_model4")

poly_model4$loadings
#help(tab_fa)

```

```{r}
sjPlot::tab_fa(poly_model4,
                title = "",
     #show.comm = TRUE,
     alternate.rows = FALSE,
     fctr.load.tlrn = 0.0)

#help(sjPlot::tab_fa)
```

```{r}
# Cluster analysis plot
fa.diagram(poly_model4)
```

## EFA considering three factors
```{r}
poly_model3 = fa(baserev, nfactor=3, cor="poly", fm="pa", rotate = "oblimin")

save(poly_model3, file = "poly_model3")

poly_model3$loadings
#help(tab_fa)

```

```{r}
sjPlot::tab_fa(poly_model3,
               title = "",
     #show.comm = TRUE,
     alternate.rows = FALSE,
     fctr.load.tlrn = 0.0)
```

```{r}
fa.diagram(poly_model3)
```

## EFA considering two factors
```{r}
poly_model2 = fa(baserev, nfactor=2, cor="poly", fm="pa", rotate = "oblimin")

save(poly_model2, file = "poly_model2")

poly_model2$loadings
#help(tab_fa)

```

```{r}
sjPlot::tab_fa(poly_model2)
```

```{r}
fa.diagram(poly_model2,
           title = "",
     #show.comm = TRUE,
     alternate.rows = FALSE,
     fctr.load.tlrn = 0.0)
```



