# Appendix

## Data manipulation

```{r}
#| label: tbl-labels
#| tbl-cap: "Change of labels in Self-efficacy items"

rbind(
data.frame(
  item=c("is2g27a","is2g27c","is2g27d","is2g27i","is2g27j","is2g27k",
         "is2g27l","is2g27m"),
  old_label=c(
  "Edit digital photographs or other graphic images",
  "Write or edit text for a school assignment",
  "Search for and find relevant information for a school project	
  on the Internet",
  "Create a multi-media presentation (with sound, pictures, or video)",
  "Upload text, images, or video to an online profile",
  "Insert an image into a document or message",
  "Install a program or [app]",
  "Judge whether you can trust information you find on the Internet"),
  new_label=c(
    "Edit graphic images",
    "Write or edit text",
    "Search and find information on internet",
    "Create multi-media presentation",
    "Upload multimedia to an online profile",
    "Insert an image into a document/message",
    "Install a program/app",
    "Judge internet information veracity")
),

data.frame(
  item=c("is2g27b","is2g27e","is2g27g","is2g27h"),
  old_label=c(
  "Create a database (e.g. using [Microsoft Access ®])",
  "Build or edit a webpage",
  "Create a computer program, macro, or [app]",
  "Set up a local area network of computers or other ICT"),
  new_label=c(
  "Create a database",
  "Build a webpage",
  "Create a computer program/app",
  "Set up a local area network")
)
) |>
  gt() |>
  tab_options(
    table.border.top.color = "white",
    heading.title.font.size = px(16),
    column_labels.border.top.width = 3,
    column_labels.border.top.color = "black",
    column_labels.border.bottom.width = 3,
    column_labels.border.bottom.color = "black",
    table_body.border.bottom.color = "black",
    table.border.bottom.color = "white",
    table.width = pct(100),
    table.background.color = "white",
    table.font.size = 10
  ) |>
  # center column text
  cols_align(align="center")|>
  tab_style(
    style = list(
      # remove horizontal lines
      cell_borders(
        sides = c("top", "bottom"),
        color = "white",
        weight = px(1)
      ),
      #center text
      cell_text(
        align="center"
      ),
    # remove row striping in Markdown documents
    cell_fill(color = "white", alpha = NULL)
      ),
      #do this for all columns and rows
    locations = cells_body(
      columns = everything(),
      rows = everything()
    )) |>
    tab_row_group(
   label = md("*Specialized Self-efficacy*"),
   rows = 9:12
     ) |>
  tab_row_group(
    label = md("*General Self-efficacy*"),
    rows = 1:8
  )|>
  cols_label(
    item=md("**Item**"),
    old_label=md("**Original label**"),
    new_label=md("**Summarised label**")
  )
```


## Test statistical significant differences

*Testing Gender differences in ICT Self-efficacy items (Chi-square)*

```{r}
# Perform chi-squared test for each item individually
chisq_test_a <- chisq.test(data$is2g27a, data$s_sex)
chisq_test_b <- chisq.test(data$is2g27b, data$s_sex)
chisq_test_c <- chisq.test(data$is2g27c, data$s_sex)
chisq_test_d <- chisq.test(data$is2g27d, data$s_sex)
chisq_test_e <- chisq.test(data$is2g27e, data$s_sex)
chisq_test_g <- chisq.test(data$is2g27g, data$s_sex)
chisq_test_h <- chisq.test(data$is2g27h, data$s_sex)
chisq_test_i <- chisq.test(data$is2g27i, data$s_sex)
chisq_test_j <- chisq.test(data$is2g27j, data$s_sex)
chisq_test_k <- chisq.test(data$is2g27k, data$s_sex)
chisq_test_l <- chisq.test(data$is2g27l, data$s_sex)
chisq_test_m <- chisq.test(data$is2g27m, data$s_sex)

# Print the results
print(chisq_test_a)
print(chisq_test_b)
print(chisq_test_c)
print(chisq_test_d)
print(chisq_test_e)
print(chisq_test_g)
print(chisq_test_h)
print(chisq_test_i)
print(chisq_test_j)
print(chisq_test_k)
print(chisq_test_l)
print(chisq_test_m)

```


*Checking normal distribution and variance homogeneity in ICT Self-efficacy indexes and CIL*

Self-efficacy indexes and CIL score gender mean differences have a normal distribution. General self-efficacy have homogeneity variance distribution, but Specialiazed self-efficacy and CIL have not a homogeneity variance distribution by gender.

```{r, echo=TRUE,results='markup'}
#| label: check-selfeff
#| tbl-cap: "Check"
shapiro.test(data$s_speceff[data$s_sex == 1])
shapiro.test(data$s_geneff[data$s_sex == 0])
shapiro.test(data$s_pv1cil[data$s_sex == 0])

leveneTest(s_geneff ~ to_label(s_sex), data = data)
leveneTest(s_speceff ~ to_label(s_sex), data = data)
leveneTest(s_pv1cil ~ to_label(s_sex), data = data)

```



## ICT Self-efficacy items correlation

```{r , echo=FALSE}
base <- student_proc_2018 |> 
  select(#s_sex,
         #s_speceff,s_geneff,
         #Items specific self-efficacy
         is2g27b,is2g27e,is2g27g,is2g27h,
         #Items general self-efficacy
         is2g27a,is2g27c,is2g27d,is2g27i,is2g27j,is2g27k,is2g27l,is2g27m)

baserev <-base |> 
       mutate_at(vars(starts_with("is2g")),
                 ~4-.)

#variables
baserev=baserev |>
  mutate(database=is2g27b,
         webpage=is2g27e,
         app=is2g27g,
         network=is2g27h,
         editphoto=is2g27a,
         edittext=is2g27c,
         findinfo=is2g27d,
         multimedia=is2g27i,
         uploadfile=is2g27j,
         insertimage=is2g27k,
         install=is2g27l,
         judgeinfo=is2g27m
         ) |>
  select(database, webpage, app, network, editphoto, edittext, findinfo, multimedia, uploadfile, insertimage, install, judgeinfo)

```

```{r, results='asis'}
#| label: tbl-continous_corr_selfeff
#| tbl-cap: Correlation matrix assuming continuous

colnames(baserev)<-c(
  "A. Create a database",
  "B. Build a webpage",
  "C. Create a computer program/app",
  "D. Set up a local area network",
  "E. Edit graphic images",
  "F. Write or edit text",
  "G. Search and find information on internet",
  "H. Create multi-media presentation",
  "I. Upload multimedia to an online profile",
  "J. Insert an image into a document/message",
  "K. Install a program/app",
  "L. Judge internet information veracity"
)

#Test correlation (Now is positive between items and index, and negativa between items and cil)
mat = round(cor(baserev, use="complete.obs"),3)
#mat


colnames(mat) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)","(H)",
                    "(I)","(J)","(K)","(L)")

#Hide lower triangle
lower<-mat
lower[lower.tri(mat, diag=TRUE)]<-""
lower<-as.data.frame(lower)

#Table using kableExtra
lower |>
  kable() |>
#kbl(caption = "Recreating booktabs style table") |>
  kable_classic()

```


```{r, results='asis'}
#| label: tbl-categorical_corr_selfeff
#| tbl-cap: Correlation matrix assuming categorical 

#Polychoric correlations
m=psych::polychoric(baserev)
polmat=m$rho

rownames(polmat)<-c(
  "A. Create a database",
  "B. Build a webpage",
  "C. Create a computer program/app",
  "D. Set up a local area network",
  "E. Edit graphic images",
  "F. Write or edit text",
  "G. Search and find information on internet",
  "H. Create multi-media presentation",
  "I. Upload multimedia to an online profile",
  "J. Insert an image into a document/message",
  "K. Install a program/app",
  "L. Judge internet information veracity"
)

colnames(polmat) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)","(H)",
                    "(I)","(J)","(K)","(L)")

#Hide lower triangle
lowerpol<-round(polmat, 3)
lowerpol[lower.tri(polmat, diag=TRUE)]<-""
lowerpol<-as.data.frame(lowerpol)


#Table using kableExtra
lowerpol |>
kable() |>
  kable_classic_2()
```

## ICT Self-efficacy items Factorial Analysis
```{r}
#| label: fig-elbow
#| fig-cap: "Paralel Analysis"

save(polmat, file = "polychoric")
  
# Factor analyses: 3 factors
load("polychoric")
fa.parallel(polmat, fm="pa", fa="fa", main = "Scree Plot", n.obs=3092)
```


```{r, results='asis'}
#| label: fig-fa4
#| fig-cap: "Factorial Analysis Self-efficacy items. 4 factors."

poly_model4 = fa(baserev, nfactor=4, cor="poly", fm="pa", rotate = "oblimin")
fa.diagram(poly_model4)
```


```{r, results='asis'}
#| label: fig-fa3
#| fig-cap: "Factorial Analysis Self-efficacy items. 3 factors."
poly_model3 = fa(baserev, nfactor=3, cor="poly", fm="pa", rotate = "oblimin")
fa.diagram(poly_model3)

```

```{r, results='asis'}
#| label: fig-fa2
#| fig-cap: "Factorial Analysis Self-efficacy items. 2 factors."

poly_model2 = fa(baserev, nfactor=2, cor="poly", fm="pa", rotate = "oblimin")
fa.diagram(poly_model2)
```

